---
title: "Data Exercise"
---

Tuberculosis (TB) remains a major global health concern, with millions of new cases reported annually. This exercise reviews and analyzes key factors associated with TB, ranging from demographic and lifestyle factors to health-related variables. The aim of analysis provides understanding of factors associated to TB.The data used in this exercise is synthetic data and was created with the help of a generative Pre-trained Transformer 3.5 (ChatGPT3.5)

```{r}
# Install and load necessary packages
#install.packages("tidyverse")
library(tidyverse)

# Set a seed for reproducibility
set.seed(123)
```

The set seed was used above for reproducibility reasons and randomly created n =998 observations

## Data Processing

```{r}
# Populate the dataframe with synthetic data
n <- 998  # Number of observations

# creating and populating Variables
tb_data <- data.frame(
  age = sample(18:65, n, replace = TRUE),  # Sample ages from 18 to 65
  gender = sample(c("Male", "Female"), n, replace = TRUE),  # Randomly assign gender
  smoking_status = sample(c("Smoker", "Non-Smoker"), n, replace = TRUE),  # Randomly assign smoking status
  diabetes_status = sample(c("Diabetic", "Non-Diabetic"), n, replace = TRUE),  # Randomly assign diabetes status
  urban_rural_status = sample(c("Urban", "Rural"), n, replace = TRUE)  # Randomly assign urban/rural status
)

# Create associations between variables and TB status
tb_data$tb_status_prob <- plogis(0.05 * (tb_data$age - 65) +
                                   ifelse(tb_data$smoking_status == "Smoker", 1, 0) +
                                   ifelse(tb_data$diabetes_status == "Diabetic", 1, 0) +
                                   ifelse(tb_data$urban_rural_status == "Urban", 0.5, 0))  # Calculate the probability of having TB

# Generate TB status based on probability
tb_data$tb_status <- ifelse(runif(n) < tb_data$tb_status_prob, 1, 0)  # Assign 1 for TB case, 0 for not TB case

# Summary statistics
summary(tb_data)

```

## Data Visualizations

Now I will carry some visualizations for some the variables

```{r}
# Boxplot of Age by TB Status
ggplot(tb_data, aes(x = factor(tb_status), y = age, fill = factor(tb_status))) +
  geom_boxplot() +
  labs(title = "Boxplot of Age by TB Status",
       x = "TB Status",
       y = "Age")

```

The 0 = No TB and 1 = TB

```{r}
# Boxplot of Age by Smoking Status
ggplot(tb_data, aes(x = factor(tb_status), y = age, fill = tb_status)) +
  geom_boxplot() +
  labs(title = "Boxplot of Age by tb_status",
       x = "tb_status",
       y = "Age")
```

```{r}
# Histogram of Age Stratified by TB Status
ggplot(tb_data, aes(x = age, fill = factor(tb_status))) +
  geom_histogram(binwidth = 5, alpha = 0.7, position = "identity") +
  labs(title = "Histogram of Age Stratified by TB Status",
       x = "Age",
       fill = "TB Status")
```

The 0 = No TB and 1 = TB

Now I will compare the distribution of age across TB Status.

```{r}
# Violin Plot of Age by TB Status
ggplot(tb_data, aes(x = factor(tb_status), y = age, fill = factor(tb_status))) +
  geom_violin() +
  labs(title = "Violin Plot of Age by TB Status",
       x = "TB Status",
       y = "Age",
       fill = "TB Status")
```

```{r}
# Bar Plot of Smoking Status Stratified by TB Status
ggplot(tb_data, aes(x = smoking_status, fill = factor(tb_status))) +
  geom_bar(position = "dodge") +
  labs(title = "Bar Plot of Smoking Status Stratified by TB Status",
       x = "Smoking Status",
       fill = "TB Status")
```

The bar graph shows that TB is more prevalent in smokers compared to the non smokers

```{r}
# Creating a table stratified by TB status

# List to store cross-tabulation tables for each variable
cross_tab_list <- list()

# Cross-tabulation table for tb_status and gender
cross_tab_gender <- table(tb_data$tb_status, tb_data$gender)
cross_tab_list[["Gender"]] <- cross_tab_gender

# Cross-tabulation table for tb_status and smoking_status
cross_tab_smoking <- table(tb_data$tb_status, tb_data$smoking_status)
cross_tab_list[["Smoking Status"]] <- cross_tab_smoking

# Cross-tabulation table for tb_status and diabetes_status
cross_tab_diabetes <- table(tb_data$tb_status, tb_data$diabetes_status)
cross_tab_list[["Diabetes Status"]] <- cross_tab_diabetes

# Cross-tabulation table for tb_status and urban_rural_status
cross_tab_urban_rural <- table(tb_data$tb_status, tb_data$urban_rural_status)
cross_tab_list[["Urban/Rural Status"]] <- cross_tab_urban_rural

# Combine all cross-tabulation tables into one table using cbind
combined_cross_tab <- do.call(cbind, cross_tab_list)

# Print the combined table
cat("Combined Cross-tabulation Table for TB Status and Other Variables:\n")
print(combined_cross_tab)
```

The above table shows the distributions of TB status across different factors

## Model Fitting

Now I will fit the different variables to look at the association using step-wise method

```{r}
# Fitting a logistic regression models 
model1 <- glm(tb_status ~ age, data = tb_data, family = "binomial")
summary(model1)
```

Model1 suggests that age is a statistically significant associated with Tuberculosis with P-value \< 0.05

```{r}
model2 <- glm(tb_status ~ age + gender,
                      data = tb_data, family = "binomial")
summary(model2)
```

Model2 suggests that gender variable may not be a statistically significant predictor of TB status in this model with p-value \> 0.05

```{r}
model3 <- glm(tb_status ~ age + gender + smoking_status,
                      data = tb_data, family = "binomial")
summary(model3)
```

Age and smoking status are statistically significant associated with Tuberculosis status in model3. However, the gender variable does not appear again to be statistically significant, as the p-value is above the conventional significance level.

```{r}
model4 <- glm(tb_status ~ age + gender + smoking_status + diabetes_status ,
                      data = tb_data, family = "binomial")
summary(model4)
```

Age, smoking status, and diabetes status are statistically significant associated with TB in this model4. However, the gender again does not appear to be statistically significant, as the p-value is above the conventional significance level.

```{r}
model5 <- glm(tb_status ~ age + gender + smoking_status + diabetes_status + urban_rural_status,
                      data = tb_data, family = "binomial")
summary(model5)
```

Age, smoking status, diabetes status, and urban/rural status are statistically significant associated with TB in model5. However,again the gender variable does not appear to be statistically significant, as the p-value is above the conventional significance level.

## Conclusion

Using Akaike Information Criterion for model selection. Model 5 has the lowest AIC (1231.2), suggesting it provides a better trade-off between goodness of fit and simplicity. The inclusion of additional variables (gender, urban/rural status) doesn't necessarily improve model fit, as indicated by AIC and residual deviance.
